---
title: "Final_Exam_653"
author: "Parvez"
format:
  html:
    embed-resources: true
editor: visual
---

# **(1) API calls to chess.com**

## Load Packages

```{r}
library(tidyverse)
library(httr)
library(jsonlite)
library(dplyr)
library(purrr)
library(tools)
```

## Player list

```{r}
response <- GET("https://api.chess.com/pub/country/US/players")

player_list <- fromJSON(rawToChar(response$content))$players

first_500 <- player_list[1:500]
```

## Self Helper Function

```{r}
get_player_data <- function(username) {
  
  info_response <- GET(paste0("https://api.chess.com/pub/player/", username))
  stats_response <- GET(paste0("https://api.chess.com/pub/player/", username, "/stats"))
  
  info <- fromJSON(rawToChar(info_response$content))
  stats <- fromJSON(rawToChar(stats_response$content))
  
  data.frame(
    username = username,
    name = ifelse(is.null(info$name), NA, info$name),
    followers = ifelse(is.null(info$followers), NA, info$followers),
    location = ifelse(is.null(info$location), NA, info$location),
    last_online = as.POSIXct(info$last_online, origin = "1970-01-01", tz = "UTC"),
    
    rapid_last_rating = ifelse(is.null(stats$chess_rapid$last$rating), NA, stats$chess_rapid$last$rating),
    rapid_best_rating = ifelse(is.null(stats$chess_rapid$best$rating), NA, stats$chess_rapid$best$rating),
    rapid_record_win = ifelse(is.null(stats$chess_rapid$record$win), NA, stats$chess_rapid$record$win),
    rapid_record_loss = ifelse(is.null(stats$chess_rapid$record$loss), NA, stats$chess_rapid$record$loss),
    rapid_record_draw = ifelse(is.null(stats$chess_rapid$record$draw), NA, stats$chess_rapid$record$draw)
  )
}
```

## Running for the first 500 players

```{r}
chess_players_df <- map_dfr(first_500, get_player_data)
```

### Viewing the data frame

```{r}
glimpse(chess_players_df)
head(chess_players_df)
```

# **(2) Identify & Remove Duplicate Files**

## R code in the Quarto document

```{r}
main_folder <- "question_2"

all_files <- list.files(
  path = main_folder,
  pattern = "\\.csv$",
  recursive = TRUE,
  full.names = TRUE
)

length(all_files)
```

## A file ID for each CSV

```{r}
file_info <- data.frame(
  file_path = all_files,
  file_hash = tools::md5sum(all_files),
  stringsAsFactors = FALSE)

head(file_info)
```

## Identifying originals

```{r}
original_files <- file_info[!duplicated(file_info$file_hash), ]

duplicate_files <- file_info[duplicated(file_info$file_hash), ]

nrow(original_files)
nrow(duplicate_files)

```

## Two-columned Table

```{r}
duplicate_table <- duplicate_files %>%
  left_join(
    original_files,
    by = "file_hash",
    suffix = c("_deleted_duplicate", "_original_retained")
  ) %>%
  select(
    original_file_retained = file_path_original_retained,
    deleted_duplicate_file = file_path_deleted_duplicate
  )

duplicate_table
```

## Deleting duplicate files

```{r}
duplicates_to_delete <- duplicate_table$deleted_duplicate_file

duplicates_to_delete <- duplicates_to_delete[file.exists(duplicates_to_delete)]

file.remove(duplicates_to_delete)
```

### Checking remaining files

```{r}
remaining_files <- list.files(
  path = "question_2",
  pattern = "\\.csv$",
  recursive = TRUE,
  full.names = TRUE
)

length(remaining_files)
```

# **(3) Multiple Models**

## Load Packages

```{r}
library(lme4)
library(knitr)

```

## Reading the Data

```{r}
benchmarks <- read_csv("benchmarks.csv")
```

## Recode season into wave

```{r}
benchmarks <- benchmarks %>%
  mutate(
    wave = case_when(
      season == "fall" ~ 0,
      season == "winter" ~ 1,
      season == "spring" ~ 2
    )
  )

table(benchmarks$season, benchmarks$wave)
```

## Function for slope reliability

```{r}
calculate_slope_reliability <- function(data, subject_name) {
  
  model_data <- data %>%
    select(sid, wave, score = all_of(subject_name)) %>%
    drop_na(score, wave)
  
  lme_model <- suppressWarnings(
    lmer(score ~ 1 + wave + (1 + wave | sid),
         REML = FALSE,
         data = model_data))
  
  true_score_variance <- as.data.frame(VarCorr(lme_model)) %>%
    filter(grp == "sid", var1 == "wave", is.na(var2)) %>%
    pull(vcov)
  
  student_slopes <- model_data %>%
    group_by(sid) %>%
    group_modify(~ {
      student_model <- lm(score ~ 1 + wave, data = .x)
      tibble(slope = coef(student_model)["wave"])
    }) %>%
    ungroup()
  
  observed_score_variance <- var(student_slopes$slope, na.rm = TRUE)
  
  slope_reliability <- true_score_variance / observed_score_variance
  
  tibble(
    subject = subject_name,
    true_score_variance = true_score_variance,
    observed_score_variance = observed_score_variance,
    slope_reliability = slope_reliability
  )}
```

## Running that function by grade and subject

```{r}
results <- benchmarks %>%
  group_by(grade) %>%
  group_modify(~ bind_rows(
    calculate_slope_reliability(.x, "reading"),
    calculate_slope_reliability(.x, "math")
  )) %>%
  ungroup()
```

## Making Table

```{r}
results %>%
  mutate(
    true_score_variance = round(true_score_variance, 3),
    observed_score_variance = round(observed_score_variance, 3),
    slope_reliability = round(slope_reliability, 3)
  ) %>%
  kable(
    caption = "Slope Reliability Estimates by Grade and Subject"
  )
```

# (4) **Slope Reliability Function**

## Function

```{r}
slope_reliability <- function(model) {
  
  model_data <- model.frame(model)
  
  outcome_name <- names(model_data)[1]
  slope_name <- names(fixef(model))[2]
  id_name <- names(lme4::getME(model, "flist"))[1]
  
  true_score_variance <- as.data.frame(VarCorr(model)) %>%
    filter(
      grp == id_name,
      var1 == slope_name,
      is.na(var2)
    ) %>%
    pull(vcov)
  
  student_slopes <- model_data %>%
    group_by(.data[[id_name]]) %>%
    group_modify(~ {
      lm_model <- lm(
        as.formula(paste(outcome_name, "~", slope_name)),
        data = .x
      )
      tibble(slope = coef(lm_model)[slope_name])
    }) %>%
    ungroup()
  
  observed_score_variance <- var(student_slopes$slope, na.rm = TRUE)
  
  true_score_variance / observed_score_variance}

```

## Creating models for Question 3

```{r}
fit_growth_model <- function(data, subject_name) {
  
  model_data <- data %>%
    select(sid, wave, score = all_of(subject_name)) %>%
    drop_na()
  
  suppressWarnings(
    lmer(
      score ~ 1 + wave + (1 + wave | sid),
      REML = FALSE,
      data = model_data))}
```

## Apply the function

```{r}
q4_results <- benchmarks %>%
  group_by(grade) %>%
  group_modify(~ {
    
    reading_model <- fit_growth_model(.x, "reading")
    math_model <- fit_growth_model(.x, "math")
    
    tibble(
      subject = c("reading", "math"),
      slope_reliability = c(
        slope_reliability(reading_model),
        slope_reliability(math_model)
      )
    )
  }) %>%
  ungroup()
```

## Simple Table

```{r}
q4_results %>%
  mutate(slope_reliability = round(slope_reliability, 3)) %>%
  kable(
    caption = "Slope Reliability Estimates from the Function"
  )

```

# **(5) 95% Confidence Interval (CI)**

## Bootstrap sample function

```{r}
boot_fun <- function(df, id_var, time_var, y_var, n_boot) {
  
  id_var <- rlang::ensym(id_var)
  
  ids <- unique(df[[rlang::as_string(id_var)]])
  n <- length(ids)
  
  boots <- vector("list", n_boot)
  
  for (i in seq_len(n_boot)) {
    
    samp_ids <- sample(ids, size = n, replace = TRUE)
    
    boot_df <- map2_dfr(
      samp_ids,
      seq_along(samp_ids),
      function(id, j) {
        df %>%
          filter(.data[[rlang::as_string(id_var)]] == id) %>%
          mutate(boot_id = paste0(id, "_", j))})
    
    boots[[i]] <- boot_df}
  
  boots}
```

## New slope reliability function

```{r}
rel_fun <- function(df, id_var, time_var, y_var) {
  
  id_name <- rlang::as_name(rlang::ensym(id_var))
  time_name <- rlang::as_name(rlang::ensym(time_var))
  
  if (is.character(y_var)) {
    y_name <- y_var
  } else {
    y_name <- rlang::as_name(rlang::ensym(y_var))
  }
  
  dat <- df %>%
    select(
      id = all_of(id_name),
      time = all_of(time_name),
      y = all_of(y_name)
    ) %>%
    drop_na()
  
  mod <- tryCatch(
    suppressWarnings(
      lmer(
        y ~ 1 + time + (1 + time | id),
        REML = FALSE,
        data = dat
      )
    ),
    error = function(e) NULL
  )
  
  if (is.null(mod)) {
    return(NA_real_)
  }
  
  true_var <- as.data.frame(VarCorr(mod)) %>%
    filter(grp == "id", var1 == "time", is.na(var2)) %>%
    pull(vcov)
  
  slopes <- dat %>%
    group_by(id) %>%
    group_modify(~ {
      lm_mod <- lm(y ~ 1 + time, data = .x)
      tibble(slope = coef(lm_mod)["time"])
    }) %>%
    ungroup()
  
  obs_var <- var(slopes$slope, na.rm = TRUE)
  
  true_var / obs_var
}


boot_grade <- function(df, subj, n_boot = 100) {
  
  boots <- boot_fun(
    df = df,
    id_var = sid,
    time_var = wave,
    y_var = subj,
    n_boot = n_boot
  )
  
  rels <- map_dbl(
    boots,
    ~ rel_fun(
      df = .x,
      id_var = boot_id,
      time_var = wave,
      y_var = subj
    )
  )
  
  tibble(
    subject = subj,
    n_size = n_distinct(df$sid),
    ci_lower = quantile(rels, 0.025, na.rm = TRUE),
    ci_upper = quantile(rels, 0.975, na.rm = TRUE),
    median_slope_reliability = median(rels, na.rm = TRUE)
  )
}
```

## Running 100 bootstraps for each grade and subject

```{r}
set.seed(123)

ci_tab <- benchmarks %>%
  group_by(grade) %>%
  group_modify(~ bind_rows(
    boot_grade(.x, "reading", n_boot = 100),
    boot_grade(.x, "math", n_boot = 100)
  )) %>%
  ungroup()

```

## Making the table

```{r}
ci_tab %>%
  mutate(
    ci_lower = round(ci_lower, 3),
    ci_upper = round(ci_upper, 3),
    median_slope_reliability = round(median_slope_reliability, 3)
  ) %>%
  kable(
    caption = "Bootstrapped 95% Confidence Intervals for Slope Reliability"
  )
```
